The impact of forest management on the temperature sensitivity of organic matter turnover

### 2024-10-23 Manuscript_draft/Draft.Rmd
Added descriptie stats about enzyme and microbial data
Fixed an error in the Ea timevariant plot (colors and sites were in the wrong order)
Small update in figures

## Parametric model
We defined a parametric model to consider all the three effects on respiration we hypothesized. The first is the influence of temperature, then the influence of soil moisture, and then a seasonal component which is assumed to be roughly proportional to the interactions of autotrophic respiration (which we hypothesized following a seasonal pattern).
Temperature and moisture scaling interact linearly, while seasonality was assumed additive:  
$$
Rh(t) =  \xi_{temp(t,j, k)} \cdot \xi_{moist(t,j)} + \xi_{sin(t,j)}
$$

@@

### Seasonality
In order to correct for eventual seasonality patters, we introduce also a seasonality scaling.
The seasonality scaling functions assume a sine function with period = 365 days, with both amplitude $ampl_j$ (the magnitude) and peak day $peak_j$ (when the function is at its only annual peak) varying as a function of treatments:
$$
\begin{aligned}
    \xi_{sin(t,k)} = ampl_k \cdot sin \left(\frac{2 \cdot \pi}{365}\right) \cdot day(t) \\
    +\left(\frac{2 \cdot \pi}{365}\right) \cdot (peak_k - 1) - \frac{\pi}{2})
\end{aligned}
$$
Both the terms $ampl_j$ and $peak_j$ are described by Bayesian priors. Variance is by treatment.
We assumed that the variability in seaslonalityis driven by treatments ($j$). 
The prior distribution for the term $ampl$ was assumed as $\sim \mathcal{N}(0, 0.5)$, so assuming as mean no seasonality effect, and for the term $peak$ was assumed as $\sim \mathcal{N}(175, 50)$, so assuming as mean for the peak the 24th of June (Midsummer day).

@@

@@

@@

## Benchmark models and parametric model fit

![Fit of the two benchmark machine learning models, linear (a) and random forest (b)](../Figures/fit_ML_validation.png){width=80%}

![Fit of the parametric model](../Figures/fit_model_validation.png){width=60%}

## Parametric model posteriors

@@

![Posteriors of the scaling parameter, varying as a function of single plots](../Figures/posteriors_boxplots_A.png){width=60%}

@@

# Appendix

@@

>>>>

## Parametric model, constant `E_a`
We defined a parametric model to consider all the three effects on respiration we hypothesized. The first is the influence of temperature, then the influence of soil moisture, and then a seasonal component which is assumed to be roughly proportional to the interactions of autotrophic respiration (which we hypothesized following a seasonal pattern).
Temperature and moisture scaling interact linearly, while seasonality was assumed additive:  
$$
Rh(t) =  \xi_{temp(t,j, k)} \cdot \xi_{moist(t,j)} + \xi_{sin(t,j)}
$$

@@

### Seasonality
In order to correct for eventual seasonality patters, we introduce also a seasonality scaling.
The seasonality scaling functions assume a sine function with period = 365 days, with both amplitude $ampl_j$ (the magnitude) and peak day $peak_j$ (when the function is at its only annual peak) varying as a function of treatments:
$$
\begin{aligned}
    \xi_{sin(t,k)} = ampl_k \cdot sin \left(\frac{2 \cdot \pi}{365}\right) \cdot day(t) \\
    +\left(\frac{2 \cdot \pi}{365}\right) \cdot (peak_k - 1) - \frac{\pi}{2})
\end{aligned}
$$
Both the terms $ampl_j$ and $peak_j$ are described by Bayesian priors. Variance is by treatment.
We assumed that the variability in seaslonalityis driven by treatments ($j$). 
The prior distribution for the term $ampl$ was assumed as $\sim \mathcal{N}(0, 0.5)$, so assuming as mean no seasonality effect, and for the term $peak$ was assumed as $\sim \mathcal{N}(175, 50)$, so assuming as mean for the peak the 24th of June (Midsummer day). Amplitude can be negative (this is just reverting the order of peaks and valleys of the sine function).

@@

## Parametric model, time-variant `E_a`
In order to test the hypothesis of an eventual variation of the activation energy of the decomposing material over time without any assumption on the function of variation, we fitted a version of the model where the $E_a$ parameter was stratified in order to let it vary independently for each treatment and year of sampling.

```{r, echo=FALSE}
# Load necessary package
library(knitr)

# Create the data frame
df <- data.frame(
  Parameter = c("A", "Ea", "a", "b", "amplitude", "peak_day"),
  Description = c(
    "Pre-exponential factor (log)", 
    "Activation energy (Ea/R)", 
    "Scaling of Moyano function (moisture, linear term)", 
    "Scaling of Moyano function (moisture, quadratic term)", 
    "Amplitude of seasonal sine wave", 
    "Peak day of seasonal sine wave"
  ),
  `Stratification Level, const Eₐ` = c(
    "Plot", 
    "Treatment", 
    "Treatment", 
    "Treatment", 
    "Treatment", 
    "Treatment"
  ),
  `Stratification Level, var Eₐ` = c(
    "Plot", 
    "Treatment-year", 
    "Treatment", 
    "Treatment", 
    "Treatment", 
    "Treatment"
  ),
    `Prior Values` = c(
    "normal(30, 100)", 
    "normal(398.5, 50)", 
    "normal(3.11, 0.1)", 
    "normal(2.42, 0.1)", 
    "normal(0, 0.25)", 
    "normal(175, 50)"
    )
)

# Generate the table using kable
kable(df, format = "pipe", caption = "Model parameters stratification and priors")
```

@@

## Machine learning variance decomposition on enzymes

@@

## Edaphic and biological variables
![Fit of the two benchmark machine learning models, linear (a) and random forest (b)](../Figures/Appendix/enzymes_boxplots.png){width=80%}

## Benchmark models and parametric model fit
![Fit of the two benchmark machine learning models, linear (a) and random forest (b)](../Figures/fit_ML_validation.png){width=80%}

![Fit of the parametric model,  constant `E_a`](../Figures/fit_model_validation.png){width=60%}

## Parametric model posteriors,  constant `E_a`

@@

## Parametric model posteriors,  time-variant `E_a` 
![Posteriors of the `E_a` parameter varying as a function of treatments and sampling year](../Figures/timeseries_Ea_timevariant.png){width=60%}

@@

# Appendix
![Posteriors of the scaling parameter, varying as a function of single plots](../Figures/posteriors_boxplots_A.png){width=60%}

@@

```{r, echo=FALSE}
# Load necessary libraries
library(knitr)

# Read the CSV file
anova_results <- read.csv("../Tables/Enzyme_means.csv", stringsAsFactors = FALSE)

# Display the table with HTML superscripts
kable(anova_results, escape = FALSE)
```
